First Exclusive Measurement of Deep Virtual Compton Scattering off $^4$He: Toward the 3D tomography of nuclei

### 2016-09-07 eg6-letters/coherent/HeliumCohDVCS.tex
Update letter

\begin{abstract}
We report the first exclusive measurement of deeply virtual Compton scatterring 
(DVCS) off a nuclei with the CLAS detector at the Jefferson laboratory. The 
observed beam spin asymmetries are significantly larger than for proton, 
confirming expectations from plane wave impulse approximation. We present an 
extraction of the unique generalized parton distribution (GPD) of the spin-0 
helium in a completely model independent way. This result demonstrate the 
strength of the method and opens a new unique access into nuclei 3D structure 
in term of quarks and gluons.
\end{abstract}

@@

The generalized parton distribution (GPD) framework developped in the past two decades
(Need a bunch of refs)~\cite{}
offers a unique access into hadron 3D structure through deeply virtual Compton
scattering (DVCS), depicted in figure~\ref{fig:DVCS} and other exclusive processes. Many measurements on the proton
at several energies have now be reported and have demonstrated the solidity of the
method~\cite{}. %To enhance a little -> CFF

Insert FIGURE DVCS nuclei
%\label{fig:DVCS}
\begin{figure}[htbp]
\caption{\label{fig:DVCS} Diagram for the coherent DVCS in the plane wave approximation.}
\includegraphics[width=8.6cm]{DVCS.png}
\end{figure}

At the same time, the partonic structure of nuclei has remained a mystery
for the past three decades, as we are still unable to explain properly the 
infamous EMC effect~\cite{}. First observed in 197X by the EMC 
collaboration, it describes the difference between the partonic structure of
free and bound nucleons. Understanding the source of this difference remains 
today a topic of active research~\cite{},
questionning our understanding of nuclei and their structure beyond protons
and neutrons~\cite{}. As studies based on inclusive measurements of the
effect seem to be too limited to tackle the question definitively, we can now
apply the GPD framework to study the nuclei in three dimensions and observe
the EMC effect in an unprecedented way.
%Elaborate on non nucleonic degrees of freedom and other interests of He DVCS

The measurement of DVCS on helium 4 is in this regard the perfect experiment for
several reasons. First, Helium 4 is of spin-0, such that a single GPD ($H$) is
necessary for the description of the DVCS at leading order. This feature allows for a
completely model independant extraction of the CFF from data that is impossible for 
protons and neutrons which need 4 GPDs, leading to complex underconstrained 
problems~\cite{}. Second, helium 4, while a small nuclei, is dense and experience 
already a significant EMC effect, while being simple enough that its nuclear
structure is very well known~\cite{} (Low energy struct, FF,...). Third, being 
light enough helium 4 can exit
a light target and be experimentally measured, insuring the exclusivity of the 
measurement. This point is especially important as we are probing with a 
multi-GeV interaction a nuclei bound by only a few MeV, this leads in the 
overwhelming majority of the cases to the incoherent break-up of the target.

%\section{Experiment}
The HERMES collaboration has measured nuclear DVCS~\cite{} without the 
possibility of separating the coherent and incoherent channels leading
to controversy in the interpretation of their results~\cite{}. We report
here our measurement of coherent DVCS off helium-4 at the Jefferson labortory
(JLab), while our results for the incoherent channel will be presented in
another paper~\cite{}. This experimental facility delivers a nearly 100\% duty factor, 6~GeV 
electron linear accelerator into three experimental halls. Our experiment
ran for 3 months in 2009 using the CEBAF Large Acceptance Spectrometer (CLAS).
This detector is covering about 2$\pi$ and is composed of drift chambers, Cherenkov counters, scintillator 
counters and an electro-magnetic calorimeter. It allows to detect accurately
electrons and provides a trigger for the experiment. It is complemented for DVCS experiments with
an inner calorimeter to detect the photons emitted at low angle and a solenoid
to keep the Moller electrons produced along the beam line to get into our detectors.

@@

In order to insure the coherence of the process, we detect all the products
of the reaction. The electron detected by the CLAS, using the drift chamber 
to reconstruct its trajectory, while the combination of the Cherenkov counters
and electromagnetic calorimeter signals insure that it is not mistaken with
a negative pion. The high energy photon is detected in a PbWO calorimeter 
placed close to the beam line. Finally, we built a small and light 
radial time projection chamber (RTPC) to 
detect the recoiling helium nuclei. The chamber is placed directly ouside of 
a gaseous target filled with 6~atm helium-4 and is 20 cm long with a 3 cm radial
drift length.

%\begin{figure}[htbp]
%\caption{\label{fig:rtpc} RTPC description}
%\includegraphics[width=8cm]{RTPC.png}
%\end{figure}

>>>>

\begin{abstract}
We report the first exclusive measurement of deeply virtual Compton scatterring 
(DVCS) off a nuclei with the CLAS detector at the Jefferson laboratory. The 
observed beam spin asymmetries are significantly larger than for proton, 
confirming expectations from calculations using the plane wave impulse 
approximation. We present an 
extraction of the single generalized parton distribution (GPD) of the spin-0 
helium-4 nuclei in a completely model independent way. This result demonstrate the 
strength of the method and opens a new and unique access into the partonic structure 
of nuclei.
\end{abstract}

@@

The generalized parton distribution (GPD) framework developped in the past 
two decades~\cite{}
offers a unique access into hadron 3D structure through deeply virtual Compton
scattering (DVCS), depicted in figure~\ref{fig:DVCS}, and other exclusive 
processes. Many measurements on the proton
at several energies have now be reported~\cite{} and phenomenological work
have demonstrated the solidity of the method to extract the tomography of the
nucleon from these data~\cite{}.

\begin{figure}[htbp]
\caption{\label{fig:DVCS} Diagram for the coherent DVCS in the plane wave approximation.}
\includegraphics[width=8.6cm]{DVCS.png}
\end{figure}

At the same time, the partonic structure of nuclei has remained a mystery
for the past three decades, as we are still unable to explain properly the 
infamous EMC effect~\cite{}. First observed in 197X by the EMC 
collaboration, it describes the difference between the partonic structure of
free and bound nucleons. Understanding the source of this difference remains 
today a topic of active research~\cite{},
questionning the basics of our understanding of nuclei and their 
structure~\cite{}. Studies based on inclusive measurements of the
effect provide a very detailed description~\cite{} but 
are still unable to tackle the question definitively. Therefore,
using the GPD framework and exclusive channels to study the nuclei in three 
dimensions will give completely new information that will affect our understanding
of the EMC effect in an unprecedented way~\cite{}. 
%In particular, the GPDs 
%allow to isolate the components of the nuclei which are not nucleonic~\cite{}.

The measurement of DVCS on helium-4 is in this regard the perfect experiment for
several reasons. First, Helium 4 is of spin-0, such that a single GPD ($H$) is
necessary for the description of the DVCS at leading order. This feature allows for a
completely model independant extraction of the CFF from data that is impossible for 
protons and neutrons which have 4 GPDs~\cite{}. Second, 
helium 4 is dense and experience 
already a significant EMC effect, while being simple enough that its nuclear
structure is well known~\cite{}. Third, being 
light enough helium-4 can exit
a light target and be experimentally measured in a coherent DVCS process, 
insuring the exclusivity of the 
measurement. This point is especially important because we are probing with a 
multi-GeV interaction a nuclei bound by only a few MeV, this leads in the 
overwhelming majority of the cases to the incoherent break-up of the target.

%\section{Experiment}
The HERMES collaboration has measured nuclear DVCS~\cite{} without the 
detection of the recoils separating the coherent and incoherent channels 
based on the measurement of the transfered 4-momentum $t$, leading
to controversy in the interpretation of their results~\cite{}. We report
here our exclusive measurement of coherent DVCS off helium-4 at the Jefferson labortory
(JLab), while our results for the incoherent channel will be presented in
another paper. JLab delivers a nearly 100\% duty factor, 6~GeV linearly 
polarized electrons into three experimental halls. Our experiment ran for 3 
months in 2009 using the CEBAF Large Acceptance Spectrometer (CLAS) in 
Hall-B. This detector is large acceptance covering nearly 2$\pi$ and is 
composed of drift chambers, Cherenkov counters, scintillator counters and 
an electro-magnetic calorimeter. It allows to detect accurately electrons 
which provide the trigger for the experiment. It is complemented for DVCS 
experiments with an inner electro-magnetic calorimeter made of PbWO crystals 
to detect the photons emitted at low angle and with a solenoid to keep the 
Moller electrons produced along the beam line to get into our detectors.

@@

In order to insure the coherence of the process, we also detect the recoiling 
helium-4 nuclei. We built a small and light radial time projection chamber 
(RTPC) to detect recoiling helium nuclei down to energies of few MeVs. The 
chamber, shown in figure~\ref{fig:rtpc}, is surrounding the gaseous target 
filled with 6~atm helium-4 and is 20 cm long with a 3 cm radial drift 
length. The detector was specifically calibrated for helium-4 nuclei using 
elastic scattering produced with a 1.2~GeV electron beam.

\begin{figure}[htbp]
\caption{\label{fig:rtpc} RTPC description}
\includegraphics[width=8cm]{RTPC.png}
\end{figure}
